### 2024-11-23 paper/al.tex
Add ambiguity section

SpecTec introduces AL, an algorithmic specification language, for specifying
the prose notation.
Given that the prose notation in the official specification is written
informally, it is essential to define AL with rigour.
We begin by how the official specification describes Wasm control flow in
\cref{sec:???}
Next, we then define a formal model for Wasm control flow in AL in
\cref{sec:model}.
We then formalize the syntax and semantics of AL in \cref{sec:formal}.
\red{Finally, we ??? in \cref{sec:spectec-prose}.}

@@

The way the official specification describe Wasm control flow differs b/w the
prose notation and the formal notation.
In the prose notation, it implicitly assumes a program counter, so it
\textit{jumps} to certain point of the code.
\begin{lstlisting}[style=wasm]
(block (br 0) (unreachable) end)
@\hatcode@
(block (br 0) (unreachable) end)
        @\hatcode@
(block (br 0) (unreachable) end)
                                  @\hatcode@
\end{lstlisting}

\begin{lstlisting}[style=wasm]
(i32.const 1) (if (br 0) (unreachable) else (unreachable) end)
@\hatcode@
(i32.const 1) (if (br 0) (unreachable) else (unreachable) end)
                @\hatcode@
(block (br 0) (unreachable) end)
@\hatcode@
(block (br 0) (unreachable) end)
        @\hatcode@
(block (br 0) (unreachable) end)
                                  @\hatcode@
\end{lstlisting}
Official specification explains Wasm control flow in a different manner in the prose notation and the formal notation.
The keyword is "jump" and "execute"

>>>>

SpecTec introduces AL, an algorithmic specification language, for specifying
the prose specification.
Since the prose specification is written informally, it is essential to define
AL with rigour.
We begin by outlining the ambiguity of the prose specification in
\cref{sec:prose-informal}.
Next, we define a formal model for Wasm control flow in AL in
\cref{sec:model}.
We then formalize the syntax and semantics of AL in \cref{sec:formal}.
\red{Finally, we ??? in \cref{sec:spectec-prose}.}

\section{Ambiguity in the prose specification}
\label{sec:prose-informal}

The prose specification describes Wasm semantics in an informal manner,
particularly when addressing control flow.
% Jump
Wasm control instructions \textit{jump} to certain code point for control
transfer.
While it does not explicitly explain the mechanics of jumping, the simplest
interpretation involves assuming a program counter.
\cref{fig:return} illustrates the prose specification for the \texttt{return}
instruction.
In the final step, it jumps to the instruction after the original call that
pushed the frame, which likely corresponds to updating the program counter to
the return address stored in the frame.

% Execute
On the other hand, some control instructions \textit{execute} a Wasm
instruction.
\cref{fig:if} represents the prose specification for the \texttt{if}
instruction, which executes a block instruction with instructions in the then
branch or the else branch based on the value in the stack.
However, the block instruction itself does not appeared in the actual code,
making its execution unclear in this context.
The situation become even more complicated when the block contains control
instructions, such as function call.
While the intended meaning can still be inferred, this informality introduces
ambiguity, leaving room for multiple interpretations.

\begin{figure}[h]
  \centerline{\includegraphics[width=15cm]{fig/return}}
  \caption[\texttt{return} instruction]{\texttt{return} instruction}
    \label{fig:return}
  \centerline{\includegraphics[width=15cm]{fig/if}}
  \caption[\texttt{if} instruction]{\texttt{if} instruction}
    \label{fig:if}
\end{figure}

@@
